---
title: "Analysis of ZnO on Si"
author: "J. R. Minter"
date: 'Started: 2018-10-17, Last modified: 2018-10-18'
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    css: ./inc/jm-gray-vignette.css
    number_sections: yes
    toc: yes
    toc_depth: 3
abstract: |
  I wanted to model EDS spectra from thin films of ZnO on Si and compute
  the K-ratios for Zn, O, and Si as a function of the thickness of the
  ZnO layer. The standards and unknowns were simulated using a python script at
  4 and 7 kV. A total of 21 unknowns were simulated at each voltage with a ZnO
  thickness ranging from 0.1 to 200 nm at each kV.
  The script performed a filter fit on each unknown using the ZnO and Si spectra
  as standards. The script then extracted the K-Ratios for the Zn La1, O Ka1,
  and Si Ka1 transitions for each unknown. DTSA-II returns these as an
  `UncertainValue2` object that reports both the mean value and the uncertainty
  of the K-ratio. These values were written to the command line and stored as a
  text file. I converted these to `.csv` files for easier processing and
  plotting in R.
  The variation of the K-ratio of each transition as a function of thickness
  was plotted for each voltage. I fit the data to a `LOESS`  (locally estimated
  scatter-plot smoothing) model and used the `predict()` function to compute a
  smooth curve in 0.1 nm increments which were written to R `tibbles` (an
  updated dataframe). These could be stored and used to estimate thickness for
  an unknown specimen.
  The 4 kV results are useful for estimation of ZnO thickness values between
  0 and 50 nm thickness with good precision. The 7 kV results are useful up to
  150 nm thickness.
---

```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
library(knitr)
library(readr)
library(pander)
library(here)
library(dplyr)
library(ggplot2)
# make this an external chunk that can be included in any file
options(width = 72)
opts_chunk$set(echo = F, message = F, warning = F)
# opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'figure/')

panderOptions('table.split.table', Inf)

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
```

# Introduction

This is a reproducible report written using RMarkdown. The accompanying
`.Rmd` file contains the code chunks that perform the computations. These
are hidden here to make the report easier to read.

Spectra were simulated using NIST DTSA-II (Kelvin, release 2018-09-26).
The simulation using 4 kV electrons was performed using the DTSA-II
python script `sim-zno-on-si-4kV.py` and the simulation using 7 kV
electrons used the script `sim-zno-on-si-4kV.py`. Each of these scripts
generated output written to the command window. This originally reported
O Ka1, ZnLa1, and SiKa1 K-ratios using the NIST `Uncertain Value2` class.
I converted these to better work as comma delimited (.csv) files.
These files were read from disk and converted to R `tibbles` (an enhanced
dataframe) that is easily plotted. The first 5 rows of the 4 kV tibble
are displayed in the table below.

```{r loadData}

fi <- here()
bSave <- FALSE

csv4kV <- '../../utility/sim-zno-on-si-4kV/sim-ZnO-on-Si-4kV.csv'
csv7kV <- '../../utility/sim-zno-on-si-7kV/sim-ZnO-on-Si-7kV.csv'

df4kV <- read_csv(csv4kV)
df7kV <- read_csv(csv7kV)
pander(head(df4kV))
```

# Compute LOESS fits

We compute LOESS  (locally estimated scatter plot smoothing) fits of the
K-ratios as a function of the ZnO thickness. R lets predict values for
intermediate points. In this case we use a smaller value for the 
`span` parameter (`0.5`) instead of the default (`0.75`). We could store
these to predict thickness values from experimental data.

```{r, computeLoessOK, results='hide'}
loess.4kV.OK <- loess(O.Ka1.KR.mu ~ t.ZnO.nm, span=0.5, data=df4kV)
# loess.4kV.OK
t.ZnO.nm <- seq(from=0.1, to=150, by=0.1)
df.lo.4kV.OK <- data.frame(t.ZnO.nm=t.ZnO.nm)
df.lo.4kV.OK$O.Ka1.KR.mu <- round(predict(loess.4kV.OK,
                                          newdata = df.lo.4kV.OK), 6)

loess.7kV.OK <- loess(O.Ka1.KR.mu ~ t.ZnO.nm, span=0.5, data=df7kV)
# loess.7kV.OK
df.lo.7kV.OK <- data.frame(t.ZnO.nm=t.ZnO.nm)
df.lo.7kV.OK$O.Ka1.KR.mu <- round(predict(loess.7kV.OK,
                                          newdata = df.lo.7kV.OK), 6)
# summary(df.lo.7kV.OK)

```


```{r, computeLoessZnLa}
loess.4kV.ZnL <- loess(Zn.La1.KR.mu ~ t.ZnO.nm, span=0.5, data=df4kV)
# loess.4kV.ZnL 
df.lo.4kV.ZnL <- data.frame(t.ZnO.nm=t.ZnO.nm)
df.lo.4kV.ZnL$Zn.La1.KR.mu <- round(predict(loess.4kV.ZnL,
                                          newdata = df.lo.4kV.ZnL), 6)

loess.7kV.ZnL <- loess(Zn.La1.KR.mu ~ t.ZnO.nm, span=0.5, data=df7kV)
# loess.7kV.ZnL
df.lo.7kV.ZnL <- data.frame(t.ZnO.nm=t.ZnO.nm)
df.lo.7kV.ZnL$Zn.La1.KR.mu <- round(predict(loess.7kV.ZnL,
                                            newdata = df.lo.7kV.ZnL), 6)
# summary(df.lo.7kV.ZnL)

```

```{r, computeLoessSiKa}
loess.4kV.SiK <- loess(Si.Ka1.KR.mu ~ t.ZnO.nm, span=0.5, data=df4kV)
# loess.4kV.SiK 
df.lo.4kV.SiK <- data.frame(t.ZnO.nm=t.ZnO.nm)
df.lo.4kV.SiK$Si.Ka1.KR.mu <- round(predict(loess.4kV.SiK,
                                             newdata = df.lo.4kV.SiK), 6)

loess.7kV.SiK <- loess(Si.Ka1.KR.mu ~ t.ZnO.nm, span=0.5, data=df7kV)
# loess.7kV.SiK
df.lo.7kV.SiK <- data.frame(t.ZnO.nm=t.ZnO.nm)
df.lo.7kV.SiK$SiK.Ka1.KR.mu <- round(predict(loess.7kV.SiK,
                                            newdata = df.lo.7kV.SiK), 6)
# summary(df.lo.7kV.SiK)

```
# Plot O K

Next we plot the O Ka1 K-ratio as a function of thickness for
both 4 and 7 kV electron beams assuming a total dose of 5000 nA-sec.

```{r plotOK}

pltO  <-  ggplot() +
          geom_point(data=df4kV, aes(x=t.ZnO.nm,
                                     y=O.Ka1.KR.mu),
                     color="darkblue") +
          geom_line(data=df.lo.4kV.OK, aes(x=t.ZnO.nm,
                                          y=O.Ka1.KR.mu),
                     color="darkblue") +
          geom_point(data=df7kV, aes(x=t.ZnO.nm,
                                     y=O.Ka1.KR.mu),
                     color="darkred") +
          geom_line(data=df.lo.7kV.OK, aes(x=t.ZnO.nm,
                                          y=O.Ka1.KR.mu),
                     color="darkred") +
          ggtitle("O Ka1 K-ratio from ZnO layers on Si with LOESS fit") +
          xlab("ZnO thickness [nm]") +
          ylab("O Ka1 K-ratio") +
          annotate("text", label = '4 kV', x = 50, y = 0.6,
                   size = 5, color = "darkblue") +
          annotate("text", label = '7 kV', x = 50, y = 0.2,
                    size = 5, color = "darkred") +
          theme(axis.text=element_text(size=12),
                axis.title=element_text(size=12),
                plot.title=element_text(hjust = 0.5)) +
        NULL
print(pltO)

if (bSave == TRUE) {
  fi <- './zno-on-si-OKa1.pdf'
  ggsave(pltO, file=fi, width=9.0, height=6.0, units="in", dpi=300)
  fi <- './zno-on-si-OKa1.png'
  # we want the png to be close to 1024x768...
  ggsave(pltO, file=fi, width=9.0, height=6.0, units="in", dpi=113.7778)
}
```

# Plot Zn La

Then we plot the Zn Ka1 K-ratio as a function of thickness for
both 4 and 7 kV electron beams.

```{r plotZnL}

pltZn <-  ggplot() +
          geom_point(data=df4kV, aes(x=t.ZnO.nm,
                                     y=Zn.La1.KR.mu),
                     color="darkblue") +
          geom_line(data=df.lo.4kV.ZnL,
                    aes(x=t.ZnO.nm, y=Zn.La1.KR.mu),
                    color="darkblue") +
          geom_point(data=df7kV, aes(x=t.ZnO.nm,
                                     y=Zn.La1.KR.mu),
                     color="darkred") +
          geom_line(data=df.lo.7kV.ZnL,
                    aes(x=t.ZnO.nm, y=Zn.La1.KR.mu),
                    color="darkred") +
          ggtitle("Zn La1 K-ratio from ZnO layers on Si") +
          xlab("ZnO thickness [nm]") +
          ylab("Zn La1 K-ratio") +
          annotate("text", label = '4 kV', x = 50, y = 0.6,
                   size = 5, color = "darkblue") +
          annotate("text", label = '7 kV', x = 50, y = 0.2,
                    size = 5, color = "darkred") +
          theme(axis.text=element_text(size=12),
                axis.title=element_text(size=12),
                plot.title=element_text(hjust = 0.5)) +
        NULL
print(pltZn)

if (bSave == TRUE) {
  fi <- './zno-on-si-ZnLa1.pdf'
  ggsave(pltZn, file=fi, width=9.0, height=6.0, units="in", dpi=300)
  fi <- './zno-on-si-ZnLa1.png'
  # we want the png to be close to 1024x768...
  ggsave(pltZn, file=fi, width=9.0, height=6.0, units="in", dpi=113.7778)
}

```

# Plot Si Ka

Then we plot the Si Ka1 K-ratio as a function of thickness for
both 4 and 7 kV electron beams.

```{r plotSiK}

pltSi <-  ggplot() +
          geom_point(data=df4kV, aes(x=t.ZnO.nm,
                                     y=Si.Ka1.KR.mu),
                     color="darkblue") +
          geom_line(data=df.lo.4kV.SiK, aes(x=t.ZnO.nm,
                                            y=Si.Ka1.KR.mu),
                    color="darkblue") +
          geom_point(data=df7kV, aes(x=t.ZnO.nm,
                                     y=Si.Ka1.KR.mu),
                     color="darkred") +
          geom_line(data=df.lo.7kV.SiK, aes(x=t.ZnO.nm,
                                           y=SiK.Ka1.KR.mu),
                    color="darkred") +
          ggtitle("Si Ka1 K-ratio from ZnO layers on Si") +
          xlab("ZnO thickness [nm]") +
          ylab("Si Ka1 K-ratio") +
          annotate("text", label = '4 kV', x = 25, y = 0.28,
                   size = 5, color = "darkblue") +
          annotate("text", label = '7 kV', x = 25, y = 0.68,
                    size = 5, color = "darkred") +
          theme(axis.text=element_text(size=12),
                axis.title=element_text(size=12),
                plot.title=element_text(hjust = 0.5)) +
        NULL
print(pltSi)

if (bSave == TRUE) {
  fi <- './zno-on-si-SiKa1.pdf'
  ggsave(pltSi, file=fi, width=9.0, height=6.0, units="in", dpi=300)
  fi <- './zno-on-si-SiKa1.png'
  # we want the png to be close to 1024x768...
  ggsave(pltSi, file=fi, width=9.0, height=6.0, units="in", dpi=113.7778)
}

```
